Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-06-10 manuscript/manuscript.tex
reorg and some cleaning
Moved images to new images folder. Changed from png to pdf format.

@@

\tableofcontents % Optional: Table of Contents
\listoffigures % List of Figures
\listoftables % List of Tables

@@

In the early 21st century, the open data movement began to take shape, driven by the 
fundamental belief that freely accessible data can transform 
both societies and governments. This movement champions the principles of 
transparency, innovation, and public engagement.

A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}.

These initiatives represent significant strides in democratizing data, in breaking barriers that once
kept valuable information on government performance in silos. Their collective impact is profound, extending beyond
mere data sharing to fostering a culture of openness that benefits
individuals, communities, governments, and economies worldwide
\citep{barns2016mine, wang2016adoption}.

@@

The contributions of this paper are twofold. First, we delve into
the specifics of data curation challenges using the NYC 311 Service
Request (SR) Data as a case study. This renowned and frequently viewed 
dataset serves as a prime example for examining key issues in data curation, 
including data validity, consistency, and curation efficiency. 
We illustrate these points with live examples drawn from our processing of the 311 SR data.

Secondly, building upon insights gained from this case study, we 
propose a set of data curation principles tailored for government-released open data. 
These principles are designed to address the unique challenges 
and requirements observed in the curation of such datasets.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Yearly.png}
  \caption{Yearly SR counts over 10-year period}
  \label{fig:yearly-counts}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{Monthly.png}
  \caption{Monthly SR counts over 10-year period}
  \label{fig:monthly-counts}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Daily.png}
  \caption{Daily SR counts over 2022-2023 period}
  \label{fig:daily-counts}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Calendar-Month.png}
  \caption{Calendar Month SR counts over a 10-year period}
  \label{fig:calendar-months-counts}
\end{figure}

@@

\begin{figure}[H]
  \centering
	  \includegraphics[width=\textwidth]{SRs_by_Agency.png}
	  \caption{SR counts by Agency and Cumulative Percentage}

@@

\begin{figure}[H]
  \centering
	  \includegraphics[width=\textwidth]{SRs_by_Complaint_Type.png}
	  \caption{Top 20 SR complaint\_type and Cumulative Percentage}
	  \label{fig:SR_complaints}
\end{figure}

@@

\begin{figure}[H]
  \centering
	  \includegraphics[width=\textwidth]{BlankFields.png}
	  \caption{Number and Percentage of Empty Entries}
	  \label{fig:blank_fields}
\end{figure}

@@

	\begin{figure}[H]
	  \centering
		  \includegraphics[width=\textwidth]{invalid_incident_zip.png}
		  \caption{Invalid incident-zip by Agency}
		  \label{fig:invalid_incident_zip}
	\end{figure}

@@

	\begin{figure}[H]
	  \centering
		  \includegraphics[width=\textwidth]{invalid_police_precincts.png}
		  \caption{Invalid NYPD precincts by Agency}
		  \label{fig:invalid_police_precincts_zip}
	\end{figure}

@@

		\begin{figure}[H]
	 	 \centering
		  \includegraphics[width = \textwidth]{invalid_community_boards.png}
		  \caption{Invalid Community Board by Agency}
		  \label{fig:invalid_community_boards}
		\end{figure}

@@

\begin{figure}[H]
 	 \centering
	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.png}
	  \caption{SRs with a Negative Duration by Agency}
	  \label{fig:negative-duration}
\end{figure}

@@

\begin{figure}[H]
 	 \centering
	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.png}
	  \caption{Negative duration distribution}
	  \label{fig:negative-duration-violin}
\end{figure}

@@

		\begin{figure}[H]
		 	 \centering
			  \includegraphics[width = \textwidth]{homeless_response_time_clean.png}
			  \caption{Homeless Assistance SR Durations}
			  \label{fig:homeless}
		\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{zero_duration_SR.png}
		 \caption{SRs with Zero Durations by Agency}
		 \label{fig:sero-duration}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.png}
		 \caption{SRs by Hourly created\_date(s)}
		 \label{fig:hourly-created}
	\end{figure}

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.png}
		 \caption{SRs by Hourly closed\_date(s)}
		 \label{fig:hourly-closed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.png}
		 \caption{SRs Created at Top of the Hour}
		 \label{fig:tophourcreated}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.png}
		 \caption{SRs created minute-by-minute on busiest day}
		 \label{fig:busiestcreated}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.png}
		 \caption{SRs Closed at Top of the Hour}
		 \label{fig:tophourclosed}
	\end{figure}

@@

 	\begin{figure}[H]
		 \centering
		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.png}
		 \caption{SRs closed minute-by-minute on busiest day}
		 \label{fig:busiestclosed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.png}
		 \caption{SRs Closed Exactly at Midnight by Agency}
		 \label{fig:midnight-closed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{closed_at_noon_chart.png}
		 \caption{SRs Closed Exactly at Noon by Agency}
		 \label{fig:noon-closed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{created_at_noon_chart.png}
		 \caption{SRs Created Exactly at Noon by Agency}
		 \label{fig:noon-created}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{post_closed_violin.png}
		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
		 \label{fig:resolution-violin}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{postClosedBarChart.png}
		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
		 \label{fig:resolution-by-agency}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.png}
		 \caption{Non-matching borough and borough\_boundary by Agency }
		 \label{fig:borough-boundaries}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.png}
		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
		 \label{fig:xstreet1}
	\end{figure}

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.png}
		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
		 \label{fig:xstreet2}
	\end{figure}

>>>>

% image path
\graphicspath{{.}{./images}}

@@

% \tableofcontents % Optional: Table of Contents
% \listoffigures % List of Figures
% \listoftables % List of Tables

@@

In the early 21st century, the open data movement began to take shape, driven by the 
fundamental belief that freely accessible data can transform 
both societies and governments. This movement champions the principles of 
transparency, innovation, and public engagement. 
A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}. 
These initiatives represent significant strides in democratizing data, in breaking barriers that once
kept valuable information on government performance in silos. Their collective impact is profound, extending beyond
mere data sharing to fostering a culture of openness that benefits
individuals, communities, governments, and economies worldwide
\citep{barns2016mine, wang2016adoption}.

@@

The contributions of this paper are twofold. First, we delve into
the specifics of data curation challenges using the NYC 311 Service
Request (SR) Data as a case study. This renowned and frequently viewed 
dataset serves as a prime example for examining key issues in data curation, 
including data validity, consistency, and curation efficiency. 
We illustrate these points with live examples drawn from our processing of the 311 SR data. 
Secondly, building upon insights gained from this case study, we 
propose a set of data curation principles tailored for government-released open data. 
These principles are designed to address the unique challenges 
and requirements observed in the curation of such datasets.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Yearly.pdf}
  \caption{Yearly SR counts over 10-year period}
  \label{fig:yearly-counts}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{Monthly.pdf}
  \caption{Monthly SR counts over 10-year period}
  \label{fig:monthly-counts}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Daily.pdf}
  \caption{Daily SR counts over 2022-2023 period}
  \label{fig:daily-counts}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{Calendar-Month.pdf}
  \caption{Calendar Month SR counts over a 10-year period}
  \label{fig:calendar-months-counts}
\end{figure}

@@

\begin{figure}[H]
  \centering
	  \includegraphics[width=\textwidth]{SRs_by_Agency.pdf}
	  \caption{SR counts by Agency and Cumulative Percentage}

@@

\begin{figure}[H]
  \centering
	  \includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf}
	  \caption{Top 20 SR complaint\_type and Cumulative Percentage}
	  \label{fig:SR_complaints}
\end{figure}

@@

\begin{figure}[H]
  \centering
	  \includegraphics[width=\textwidth]{BlankFields.pdf}
	  \caption{Number and Percentage of Empty Entries}
	  \label{fig:blank_fields}
\end{figure}

@@

	\begin{figure}[H]
	  \centering
		  \includegraphics[width=\textwidth]{invalid_incident_zip.pdf}
		  \caption{Invalid incident-zip by Agency}
		  \label{fig:invalid_incident_zip}
	\end{figure}

@@

	\begin{figure}[H]
	  \centering
		  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}
		  \caption{Invalid NYPD precincts by Agency}
		  \label{fig:invalid_police_precincts_zip}
	\end{figure}

@@

		\begin{figure}[H]
	 	 \centering
		  \includegraphics[width = \textwidth]{invalid_community_boards.pdf}
		  \caption{Invalid Community Board by Agency}
		  \label{fig:invalid_community_boards}
		\end{figure}

@@

\begin{figure}[H]
 	 \centering
	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
	  \caption{SRs with a Negative Duration by Agency}
	  \label{fig:negative-duration}
\end{figure}

@@

\begin{figure}[H]
 	 \centering
	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
	  \caption{Negative duration distribution}
	  \label{fig:negative-duration-violin}
\end{figure}

@@

		\begin{figure}[H]
		 	 \centering
			  \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
			  \caption{Homeless Assistance SR Durations}
			  \label{fig:homeless}
		\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
		 \caption{SRs with Zero Durations by Agency}
		 \label{fig:sero-duration}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.pdf}
		 \caption{SRs by Hourly created\_date(s)}
		 \label{fig:hourly-created}
	\end{figure}

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.pdf}
		 \caption{SRs by Hourly closed\_date(s)}
		 \label{fig:hourly-closed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.pdf}
		 \caption{SRs Created at Top of the Hour}
		 \label{fig:tophourcreated}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.pdf}
		 \caption{SRs created minute-by-minute on busiest day}
		 \label{fig:busiestcreated}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.pdf}
		 \caption{SRs Closed at Top of the Hour}
		 \label{fig:tophourclosed}
	\end{figure}

@@

 	\begin{figure}[H]
		 \centering
		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.pdf}
		 \caption{SRs closed minute-by-minute on busiest day}
		 \label{fig:busiestclosed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
		 \caption{SRs Closed Exactly at Midnight by Agency}
		 \label{fig:midnight-closed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{closed_at_noon_chart.pdf}
		 \caption{SRs Closed Exactly at Noon by Agency}
		 \label{fig:noon-closed}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{created_at_noon_chart.pdf}
		 \caption{SRs Created Exactly at Noon by Agency}
		 \label{fig:noon-created}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{post_closed_violin.pdf}
		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
		 \label{fig:resolution-violin}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{postClosedBarChart.pdf}
		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
		 \label{fig:resolution-by-agency}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.pdf}
		 \caption{Non-matching borough and borough\_boundary by Agency }
		 \label{fig:borough-boundaries}
	\end{figure}

@@

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.pdf}
		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
		 \label{fig:xstreet1}
	\end{figure}

	\begin{figure}[H]
		 \centering
		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.pdf}
		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
		 \label{fig:xstreet2}
	\end{figure}
